---
title: "Regression HW3"
output: html_notebook
---
```{r}
library(tidymodels)
```

1

```{r}
df <- read.table(file="P083.txt", header = TRUE, sep ="\t")
print(df)
```

```{r}
#3.3a
print("3.3a")
lm(F ~ P1, data = df) %>% 
  summary()
```
```{r}
lm(F ~ P2, data = df) %>% 
  summary()
```

```{r}
lm(F ~ P2 + P1, data = df) %>% 
  summary()
```
3.3b)
```{r}
t1 = -22.3424/11.5640
t2 = -1.85355/7.56181
t3 = -14.5005/9.2356
#F3
# aov(F ~ P2+P1, data = df) %>%
#   summary()
# anova(rms::ols(F ~ P2, data = df))
```

```{r}
df <- read.table(file="P060.txt", header = TRUE, sep ="\t")
print(df)
```
```{r}
#FM
anova(rms::ols(Y ~ X1 +X2 +X3, data = df)) 
```

```{r}
lm(Y ~ X1 +X3, data = df) %>% 
  summary()
```

```{r}
#RM
anova(rms::ols(Y ~ X1 +X3, data = df)) 
```

```{r}
df <- read.table(file="P088.txt", header = TRUE, sep ="\t")
print(df)
```
```{r}
lm(Sales~Age+HS+Income+Black+Female+Price, data = df) %>%
  summary()
```

```{r}
anova(rms::ols(Sales~Age+HS+Income+Black+Female+Price, data = df))
```

```{r}
anova(rms::ols(Sales~Age+Income+Black+Price, data = df))

```

```{r}
lm(Sales~Age+HS+Black+Female+Price, data = df) %>%
  summary()
```
```{r}
lm(Sales~Age+Income+Price, data = df) %>%
  summary()
```

```{r}
lm(Sales~Income, data = df) %>%
  summary()
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
