---
title: "R HW01"
output:
  html_document:
    df_print: paged
---

Load libraries
```{r}
library(tidyverse)
```
1. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg? Include your code please.

A: The categorical variables are: manufacturer, model, trans, drv, fl, class, model
The continuous variables: displ
It is worth mentioning that although hwy and cty are in integer format, they can actually be in decimal format (which would make them continuous).  
```{r}
glimpse(mpg)
?mpg
```
2. What happens if you map the same variable to multiple aesthetics? Include your code please.


A: When you map the same variable to multiple aesthetics, all of the aesthetics are dependent on the variable.
```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(colour = class, alpha=class, shape=class, fill=class))
```

When you map the same graph with multiple aesthetics, there are 2 scenarios. 
1. You pass different parameters to the aesthetics (ex. one is aes(color=class) and the other is aes(x=displ, y=hwy)) 
The plot accepts both aesthetic settings and combines them to become one.
2. You pass the same parameter to the aesthetics. 
Either each plot follows its own aesthetic (for example you have text and point different color, they won't be both the same color) or the entire plot gets affected and the original parameters are overriden.
```{r}
#Case 1
#Just 1 aesthetic
ggplot(mpg) +
 geom_point(aes(x=displ, y =hwy, colour = class))
#2 aesthetics
ggplot(mpg, aes(x = displ, y = hwy)) +
    geom_point(aes(colour = class))

```
```{r}
#Case 2 
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)
#Multiple aesthetics with no over-riden
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(colour = class)) +
  geom_text(aes(color="black", label = model), data = best_in_class)
#Multiple aesthetics with override
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(colour = class)) +
  geom_text(aes(x=hwy, y=displ,color=class, label = model), data = best_in_class)
```


3. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y. Include your code please.

The aesthetic (in this case the color) is separated by those that meet the logical argument and those that don't. 

```{r}
ggplot(mpg) +
 geom_point(aes(x=displ, y =hwy, colour = displ <5 ))
ggplot(mpg) +
 geom_smooth(aes(x=displ, y =hwy, colour = class == "compact" ))

```

4. What happens if you facet on a continuous variable? Include your example code please.

A: If one facets on a continuous variable, there will be a plot for every continuous value. This is really inefficient and should be considered bad practice. 
```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
geom_point(aes(colour = class)) + 
facet_wrap(~displ)  

```

5. What plots does the following code make? What does . do? Include your code please.

A:The "." changes the orientation of the graph. If it is placed after the ~, then the facet is horizontally-oriented, if it is placed before the ~, then it will be vertically arranged. 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ drv)
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(cyl ~ .)
```

6. Recreate the R code necessary to generate the following graphs. [last plot is bonus]
```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
geom_point() +
geom_smooth(se=FALSE) 
```
```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
geom_point() +
geom_smooth(mapping = aes(group=drv), se=FALSE) 
```
```{r}
ggplot(mpg, aes(x = displ, y = hwy, color=drv)) +
geom_point() +
geom_smooth(mapping = aes(group=drv), se=FALSE) 
```
```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
geom_point(aes(color=drv)) +
geom_smooth(se=FALSE) 
```
```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
geom_point(aes(color=drv)) +
geom_smooth(mapping = aes(group=drv, linetype=drv), se=FALSE) 
```
```{r}
ggplot(mpg, aes(x = displ, y = hwy, fill=drv)) +
geom_point(shape=21, color ="white", stroke=4, size=5)
```


7. What is the problem with this plot? How could you improve it?

A: There's no geom_plot, smooth, or bar, or anything. Therefore the graph results in nothing. One needs to choose what type of visualization to do in order to visualise anything. 
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy))
ggplot(data = mpg, mapping = aes(x = cty)) + geom_bar()
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + geom_smooth()
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + geom_line() 
```