---
title: "STA205 MIDTERM EXAM - Ian Liu"
author: "Ian Liu"
date: 'Due: 11pm Saturday March 12, 2022'
output:
  html_notebook:
    fig_height: 4
    fig_width: 7
  html_document:
    df_print: paged
  pdf_document: default
---



## Academic Honesty Statement

I, Ian Liu, hereby state that I have not communicated with or gained information in any way from my classmates or anyone other than the Professor during this exam, and that all work is my own.

## Load packages

```{r load-packages, message=FALSE}
# load required packages here
library(tidyverse)
library(nycflights13)
```

## Questions 

### Question 1: 

```{r}
group_by(flights, dest) %>%
    dplyr::summarize(n=n()) %>% 
    arrange(desc(n))
```

As you can see above, the 10 most common destinations (ordered from having most flight destinations to least) for flights departing from NYC in 2013 are: Chicago O'Hare International Airport (ORD), Hartsfield-Jackson Atlanta International Airport, Los Angeles International Airport, Boston Logan International Airport, Orlando International Airport, Charlotte Douglas International Airport, San Francisco International Airport, Fort Lauderdale-Hollywood International Airport, Miami International Airport, Ronald Reagan Washington National Airport. 

### Question 2

```{r}
group_by(flights, carrier) %>%
    dplyr::summarize(n=n()) %>% 
    arrange(desc(n))
```

Now that we have the carrier code for the top 10 airlines, we need to find their actual airline names. 

```{r}
airlines
```

The top 10 airlines with the most flights departing from NYC in 2013 are (from most flights to least): United Air Lines Inc., JetBlue Airways, ExpressJet Airlines Inc., Delta Air Lines Inc., American Airlines Inc., Envoy Air, US Airways Inc., Endeavor Air Inc., Southwest Airlines Co., Virgin America.

### Question 3

```{r}
filter(flights, !is.na(arr_delay)) %>%
    group_by(carrier) %>% 
    dplyr::summarize(n=n(), mean_arr_delay=mean(arr_delay),    
                      sum_arr_delay=sum(arr_delay)) %>%
    arrange(desc(mean_arr_delay))

# The reason I have count and sum is to verify that the mean is indeed correct.
```

3a) The carrier with the highest mean arrival delay is Frontier Airlines Inc. (F9).

3b) The carrier with the lowest mean arrival delay is Alaska Airlines Inc. (AS).

### Question 4

```{r}
#First let's find the most delayed flight in the data set, its date, and its origin. 
arrange(flights, desc(dep_delay)) %>% 
    select(year, month, day, dep_delay, origin)
```

```{r}
#Now that we have its origin and date, we can find out what the temperature was like on that day using the weather dataset. (Here we use a simple, no-join method, but joins can also do the same.)
filter(weather, origin=="JFK", month==1, day==9) %>%
    dplyr::summarise(mean_temp=mean(temp))

```
On January 9th, the day with the highest departure delay of 2013, the mean temperature at JFK, the location of the departure delay, was 42.6575 degrees Fahrenheit. 

### Question 5

```{r}
#First we find the flights that have non-null arrival delay times and create a new categorical variable called arrival, which tells  us if it was on-time or delayed. 
flights_w_arrival = filter(flights, !is.na(arr_delay)) %>%
  mutate(arrival = ifelse(arr_delay <=0, "ontime", "delayed"))

#Next, since we only want the flights with destination to PHL and RDU, we will filter out all the other values. We will also filter out flights that aren't in December.
phl_rdu_flightswarr = filter(flights_w_arrival, dest == "PHL" | dest == "RDU", month== 12)

```

```{r}
ggplot(data=phl_rdu_flightswarr, mapping=aes(y=dep_delay, x=arrival)) +
    geom_boxplot(aes(group=arrival, color=arrival)) +
    facet_grid(dest ~ origin) +
    labs(title="On time performance of NYC flights December 2013")
```

I think the purpose of this visualization is to compare the ratio of December flights that were ontime and delayed. We also want to see if there's a particular airport that has more delays or which destination may cause more delays. 

Since we're comparing with the month of December, we may also want to see if snow has an effect on these 2 airports. After understanding which origin or which destination may cause more delays, we will want to do further investigation in the exploratory data analysis. 

### Extra Credit


```{r}
flights_by_day_and_origin = filter(flights, !is.na(dep_delay)) %>%
    group_by(month,day, origin) %>%
    transmute(avg_daily_dep_delay = mean(dep_delay), month=month,day=day,origin=origin)
temp_by_day_and_origin = filter(weather, !is.na(temp)) %>%
    group_by(month,day, origin) %>%
    transmute(avg_daily_temp=mean(temp), month=month,day=day,origin=origin)

FlightAndTempData = inner_join(x=flights_by_day_and_origin, y=temp_by_day_and_origin, 
           by=c("month" = "month", "day" = "day", "origin" = "origin"))

```

```{r}
ggplot(data=FlightAndTempData, mapping=aes(x=avg_daily_dep_delay, y=avg_daily_temp)) + 
  geom_point(mapping=aes(color=origin))
```
```{r}
ggplot(data=FlightAndTempData, mapping=aes(x=avg_daily_dep_delay, y=avg_daily_temp)) + 
  geom_line(mapping=aes(color=origin))
```

